A Hierarchical Bayesian Framework for Inferring the Stellar Obliquity Distribution

### 2023-04-13 src/tex/ms.tex
[showyourwork] overleaf sync

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system. In transiting exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured. To determine the true stellar obliquity $\psi$, information about the stellar inclination $i_\star$ is also needed.
In this paper, we demonstrate that while a constraint on the stellar inclination is crucial for measuring the obliquity of an individual system, it is not required for robust determination of the population-level stellar obliquity distribution. 
In practice, the constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
We introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available.
When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees. Misaligned systems have nearly isotropic stellar obliquities \edits{with no strong clustering near 90$\degr$. 
Furthermore, misaligned Hot Jupiters have a diverse range of stellar obliquities, suggesting that dynamic mechanisms, such as planet-planet scattering after a convergent disk migration or planetary Kozai, are likely responsible for producing the population.}

@@

\begin{figure*}[ht!]
    \script{transform.py}
    \includegraphics[width=\linewidth]{figures/transform.pdf}
    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column). The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column. The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
    \label{fig:transform}
\end{figure*}

>>>>

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system. In transiting exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured. To determine the true stellar obliquity $\psi$, information about the stellar inclination $i_\star$ is also needed.
In this paper, we demonstrate that while a constraint on the stellar inclination is crucial for measuring the obliquity of an individual system, it is not required for robust determination of the population-level stellar obliquity distribution. 
In practice, the constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
We introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available.
When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees. Misaligned systems have nearly isotropic stellar obliquities \edits{with no strong clustering near 90$\degr$. 
Furthermore, misaligned Hot Jupiters have a diverse range of stellar obliquities from $0\degr$ to $180\degr$, preferring dynamic mechanisms, such as planet-planet scattering after a convergent disk migration or planetary Kozai, which could produce broad distributions.}

@@

\begin{figure*}[ht!]
    \centering
    \script{transform.py}
    \includegraphics[width=0.8\linewidth]{figures/transform.pdf}
    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column). The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column. The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
    \label{fig:transform}
\end{figure*}
